### 2022-11-29 notes.tex
Lots of progress on rust section

% QUESTIONS:
% - difference between aliases and references
% - difference between "a is owner of b" and "b is outside a"
% - how to reference that a whole paragraph came from one source?
% - how to incorporate tables? What tables could be relevant in my paper?
% - Online article: Should I cite the actual pdf or the website that brought me to the pdf
% - british english vs american english spelling of same words?
% - should i really call it "Main part"? Or split up main part into own sections?
% - Is what i have now on owners-as-... what should go in state of the art?

\documentclass[sigplan,11pt,nonacm]{acmart}
\settopmatter{printfolios}

\usepackage{booktabs} % For formal tables
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{hyphenat}
\usepackage{todonotes}
\usepackage[babel]{csquotes}
\usepackage{listings}
\lstset{frame=tb,
  language=Rust,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true,
  tabsize=3
}

\begin{document}
\title{Ownership types in theory and practice (in Rust)}
\author{Fritz Rehde}
\affiliation{%
  \institution{Technical University of Munich}
}
\email{fritz.rehde@tum.de}

\begin{abstract}
% Abstract: Brief summary of area, problem, approach, key result

- why do we need ownership types, what do they do?

\end{abstract}

\keywords{Ownership, Type, Safety, Rust}

\maketitle

\section{Introduction}
\label{sec:introduction}
% Introduction: introduce area, problem, approach, key results, contributions, outline

\section{Background}
\label{sec:background}
% Background: if needed, describe prerequisites

\subsection{Memory Safety}

source: Definition of memory safety\cite{memory1}.
source: Why memory safery is important\cite{nsa-memory-safety}.

\subsection{State of the art}

- when/in which language first introduced?
- why slow adoption?

 - TODO: source: Standard ML, C++?

- source: \cite{understanding-evolving-rust}
- source: \cite{ownership-types-survey}
Notes from Ownership Types Survey\cite{ownership-types-survey}

before: simply protect fields of an object from external access
with OT: also protect objects stored in those fields by claiming ownership of and access to other objects
problem OTs solve: object aliasing in OOP
bugs due to unintentional aliases, difficult to find
potential solutions: banning aliases, clearly advertising aliases, manage and control its effects
page 16:
1998: flexible alias protection:
limit visibility of changes to objects via aliases
by limiting where alias could propagate
by limiting the changes that could be observed through alias
didn't feature "this"
evolved into original OT:

owners-as-dominators invariant:
- heap divided into next regions (ownership contexts)
- OC is set of objects
- organised hierarchically: each owned object inside its owner, nesting relationship
- each object has one single context
- receives permissions to reference objects in other contexts through type

core concept: object ownership: objects owned by other object or entities (global owners, stack-based owners)
objects in same ownership context: siblings or peers

control aliasing by removing aliasing
unique references: only reference to object in system
borrowing: 
- reference to owned object temporarily passed to another object (for method call)
- method over => all temporary references vanish/unusable

control aliasing by controlling effects of aliasing
immutability: state cannot change

owners-as-dominators
- program heap is tree structured, object is inside its owner
- external access only through interface, i.e. owner
- rules for a to validly access b: 1. a is owner of b; 2. a and b are siblings; 3. b is outside a
- "a is inside b": "a owned by b": a = b or a is owned by b, transitive
- "a is outside b": converse of inside: "b is inside a"
- advantages: simple, clear and strong garantee, easy to reason about properties of code
- disadvantages: no aliasing => no common idioms; programming more difficult

owners-as-modifiers
- weaker form of owners-as-dominators, allows read-only references
- read-only reference: only used to read fields, call pure methods (may not modify any existing object, including receiver)
- aliasing unrestricted, but only owner can modify
- usecase: requirements from verification of object-oriented programs
- rules for a to validly reference b through reference r: either
1. a is owner of b
2. a and b are siblings
3. b is outside a
4. r is read-only reference and only pure methods can be called

ownership domains:

owners-as-ombudsmen:
- problem: strong topological requirement: tree structure: some idioms don't fit in tree structure
- allow multiple objects to define a shared aggregate owner
- rules for a to validly reference b through reference r: either
1. a is owner of b
2. a and b are siblings
3. b is outside a
4. a is owned by aggregate owner b

\section{Main part}
\label{sec:mainpart}
% Main part (approach, evaluation, discussion, etc.)

\subsection{Ownership types in theory}

\subsection{Implementation in Rust}
- borrow checker

Notes from rust-book\cite{rust-book}.
Ownership definition: set of rules that govern how Rust program manages memory
Compiler checks rules, if violated program doesn't compile
Purpose of Rust ownership is to mainly manage heap data
TODO: contradiction to below: value can be invalid in scope
Ownership rules:
- each value in Rust has an owner
- there can only be one owner at a time
- when the owner goes out of scope, the value will be dropped (value is valid in scope, invalid out of scope)
Definition scope:

\begin{lstlisting}
{                      // s is not valid here, it’s not yet declared
    let s = "hello";   // s is valid from this point forward
    // do stuff with s
}                      // this scope is now over, and s is no longer valid
\end{lstlisting}

data types of known size at compile time (int, float, bool, char): stored on stack, popped off stack when out of scope, quickly and trivially copied
heap: mutable data
stack: immutable data, size known at compile time
requirements of memory allocation system:
1. memory must be requested from memory allocator at runtime
2. need way of returning/freeing memory when done using it

First part is trivial
Second part: other solutions:
1. garbage collector: keeps track of and cleans up memory isn't used anymore (we don't need to think about it/do anything)
2. without GC, traditional: explicitly free memory
- difficult
- if forget, waste memory
- if too early, invalid variable
- if free twice, bug
=> need to pair one allocate with exactly one free
Rust: memory is automatically freed once variable that owns it goes out of scope
Rust calls `drop` automatically at end of every scope

Move:
simple (find different word) data that is stored on stack: data is copied
example:
\begin{lstlisting}
let x: i32 = 42;
let y: i32 = x;
\end{lstlisting}

For datastructures stored on the heap like String:

\begin{lstlisting}
let x: String = String::from("hello world");
let y: String = x;
\end{lstlisting}

similar to shallow copy of fields stored on stack (pointer to memory in heap, maybe length or capacity as well).
Assumption: both x and y point to the same memory in the heap.
what happens when they go out of scope, what is freed?
One object cannot have two owners.
In rust: variable x is moved into y = y becomes the new owner of the object and x is invalid

% - TODO: more sources

\subsection{Practical examples using Rust}
% - show common Rust ownership errors
% - Protection against memory vulnerabilitie examples
% - reference: \cite{rust-by-example}

\subsection{Alternatives}
% - comparison to alternative type systems without ownership (C/C++: none at all, Java: garbage collection)
% - choose non-ownership programming language e.g. C++20
% - compare code examples
% - compile time differences between Rust and non-ownership lang
% - runtime differences between Rust and non-ownership lang

\section{Related work}
\label{sec:relatedwork}
% (In a paper: Related Work – might come before main part)

\section{Summary \& Outlook}
\label{sec:summary}

% QUESTIONS:
% Online article: Should I cite the actual pdf or the website that brought me to the pdf

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper} % read paper.bib file

\end{document}