### 2023-01-13 slides/slide-template.tex
Restructured repo

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{tikz}
\usepackage[style=alphabetic,giveninits=true,terseinits=true]{biblatex}

\definecolor{tumblue}{RGB}{0,101,189}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\lstset{
    basicstyle=\footnotesize\ttfamily,%
    columns=fullflexible,%
    keywordstyle=\bfseries\color{blue},%
    commentstyle=\color{dkgreen},%
    stringstyle=\color{mauve},%
}

\DeclareFieldFormat{url}{\href{#1}{\resizebox{!}{2ex}{\lower3pt\hbox{\pgfuseimage{beamericononline}}}}}
\addbibresource{slides.bib}

\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\usecolortheme{orchid}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{sidebar right}{% also implies no navbar
  \llap{\tikz\fill[tumblue,scale=.1] (0,0) ++(-5cm,-5cm) ++(-10cm,0)
      +(0cm,0cm) -- +(4cm,0cm) -- +(4cm,-4cm) -- +(5cm,-4cm) -- +(5cm,0cm) --
      +(10cm,0cm) -- +(10cm,-5cm) -- +(9cm,-5cm) -- +(9cm,-1cm) --
      +(8cm,-1cm) -- +(8cm,-5cm) -- +(7cm,-5cm) -- +(7cm,-1cm) --
      +(6cm,-1cm) -- +(6cm,-5cm) -- +(3cm,-5cm) -- +(3cm,-1cm) --
      +(2cm,-1cm) -- +(2cm,-5cm) -- +(1cm,-5cm) -- +(1cm,-1cm) --
      +(0cm,-1cm) -- cycle;}
  % Slide number in triangle.
  \vfill\llap{\tikz\path[fill=tumblue,inner sep=3pt] (0,0) -- ++(-1cm,0) -- +(1cm,1cm) -- cycle node[anchor=south east] {\usebeamerfont{footline}\color{white}\bfseries\insertframenumber};}%
  % Variant without triangle
  %\vfill\llap{\usebeamerfont{footline}\usebeamercolor[fg]{footline}\insertframenumber\hskip3pt}\vskip3pt%
}

\title{Nice Presentation Title}
\author{Your Name \and Your Co-author's Name}
\institute{School of Computation, Information, and Technology\\Technical University of Munich}
\date{Future}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Organization}
\begin{itemize}
  % Don't forget to adjust the \addbibresource above.
  \item Citation example, if you really need it\footfullcite{mysource}
\end{itemize}
\end{frame}

\end{document}