### 2023-01-22 slides/slides.tex
Some progress on slides

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}
% \usepackage[style=alphabetic,giveninits=true,terseinits=true]{biblatex}

\definecolor{tumblue}{RGB}{0,101,189}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{mauve}{rgb}{0.58,0,0.82}
% \lstset{
%   frame=tb,
%   basicstyle=\footnotesize\ttfamily,
%   columns=fullflexible,
%   numbers=left,
%   numbersep=5pt,
%   xleftmargin=\parindent,
%   keywordstyle=\bfseries\color{blue},
%   commentstyle=\color{dkgreen},
%   stringstyle=\color{mauve},
% }

\lstset{
  frame=tb,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle=\footnotesize\ttfamily,
  % basicstyle={\small\ttfamily},
  numbers=left,
  numbersep=5pt,
  % xleftmargin=\parindent,
  numberstyle=\color{gray},
  keywordstyle=\bfseries\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=false,
  tabsize=2
}
\definecolor{cgreen}{RGB}{60,201,65}

% \DeclareFieldFormat{url}{\href{#1}{\resizebox{!}{2ex}{\lower3pt\hbox{\pgfuseimage{beamericononline}}}}}
% \addbibresource{slides.bib}

\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\usecolortheme{orchid}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{sidebar right}{% also implies no navbar
  \llap{\tikz\fill[tumblue,scale=.1] (0,0) ++(-5cm,-5cm) ++(-10cm,0)
      +(0cm,0cm) -- +(4cm,0cm) -- +(4cm,-4cm) -- +(5cm,-4cm) -- +(5cm,0cm) --
      +(10cm,0cm) -- +(10cm,-5cm) -- +(9cm,-5cm) -- +(9cm,-1cm) --
      +(8cm,-1cm) -- +(8cm,-5cm) -- +(7cm,-5cm) -- +(7cm,-1cm) --
      +(6cm,-1cm) -- +(6cm,-5cm) -- +(3cm,-5cm) -- +(3cm,-1cm) --
      +(2cm,-1cm) -- +(2cm,-5cm) -- +(1cm,-5cm) -- +(1cm,-1cm) --
      +(0cm,-1cm) -- cycle;}
  % Slide number in triangle.
  \vfill\llap{\tikz\path[fill=tumblue,inner sep=3pt] (0,0) -- ++(-1cm,0) -- +(1cm,1cm) -- cycle node[anchor=south east] {\usebeamerfont{footline}\color{white}\bfseries\insertframenumber};}%
  % Variant without triangle
  % \vfill\llap{\usebeamerfont{footline}\usebeamercolor[fg]{footline}\insertframenumber\hskip3pt}\vskip3pt%
}

\title{Ownership types in theory and practice (in Rust)}
\author{Fritz Rehde}
\institute{School of Computation, Information, and Technology\\Technical University of Munich}
\date{26.01.2023}

% \useoutertheme[subsection=false]{miniframes}

\begin{document}

\frame{\titlepage}

\begin{frame}{Overview}
\tableofcontents
\end{frame}

\section{Motivation}

\begin{frame}{Aliasing}
\begin{itemize}
  \item object aliasing = accessing the same memory through different symbolic names
  % TODO: example
  \item unintentional aliases $\rightarrow$ bugs, hard to detect
  \item ownership types goal: restrict object aliasing (through rules) $\rightarrow$ improve memory safety
\end{itemize}
\end{frame}

\begin{frame}{Memory safety}
\begin{itemize}
  \item ownership types goal: improve memory safety
  \item Google and Microsoft: 70 percent of vulnerabilities due to memory safety issues
  \item National Security Agency (USA) recommends memory-safe languages (instead of C/C++)
  \item examples: buffer overflows, memory leaks, double-free, use-after-free
  \item result: malicious exploits, incorrect program results, "random" crashes
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Memory safety: example}
C++
\begin{lstlisting}[language=C++]
std::vector<int> v { 10, 11 };
int *vptr = &v[1];  // points into v
v.push_back(12);    // v buffer reallocated => vptr dangling
std::cout << *vptr; // bug (use-after-free)
\end{lstlisting}

Rust
\begin{lstlisting}[language=Rust]
let mut v = vec![10, 11];
let vptr = &mut v[1];  // 1. mutable reference
Vec::push(&mut v, 12); // 2. mutable reference
println!("{}", *vptr); // compiler error
\end{lstlisting}
\end{frame}

\section{Ownership types in theory}

\begin{frame}{Ownership types}
\begin{itemize}
  \item object-oriented programs:
  \begin{itemize}
    \item objects can reference any other object
    \item objects can read and modify other objects' fields
    \item makes it difficult to understand, to maintain, and to reason about
  \end{itemize}
  \item using ownership types:
  \begin{itemize}
    \item limit which objects can be referenced
    \item can specify whether the referenced objects may be mutated or just read from
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Ownership types: core concepts}
\begin{itemize}
  % TODO: different highlighting than italics
  \item objects can be \emph{owners} of other objects
  \item \emph{inside} relation: \emph{a} is \emph{inside} \emph{b} if they are the same object or \emph{a} is transitively owned by \emph{b}
  \item \emph{outside} relation: converse of \emph{inside} relation
  \item \emph{ownership context}: set of all objects that an object owns
  \item \emph{siblings}: objects in the same ownership context
  \item special \emph{world} object is root of the ownership hierarchy tree
\end{itemize}

% TODO: make spacing between vertical objects smaller
\begin{figure}[fragile]
  \begin{tikzpicture}[object/.style={rectangle, draw, very thick, minimum size=20}]
    \node[object] (World) {world};
    \node[object] (Car) [below=of World] {car};
    \node[object] (Driver) [right=of Car] {driver};
    \node[object] (Engine) [below=of Car] {engine};

    \draw[->,thick] (World) to (Car);
    \draw[->,thick] (World) to (Driver);
    \draw[->,thick] (Car) to (Engine);
  \end{tikzpicture}

  \caption{solid lines indicate "owns"}
\end{figure}
\end{frame}

\begin{frame}{Owners-as-dominators}
For \emph{a} to validly reference \emph{b}:
\begin{enumerate}
  \item \emph{a} is the owner of \emph{b},
  \item \emph{a} and \emph{b} are siblings, or
  \item \emph{b} is outside of \emph{a}.
\end{enumerate}

\begin{figure}[fragile]
  \begin{tikzpicture}[object/.style={rectangle, draw, very thick, minimum size=20}]
    \node[object] (World) {world};
    \node[object] (Car) [below=of World] {car};
    \node[object] (Driver) [right=of Car,xshift=1.5cm] {driver};
    \node[object] (Engine) [below=of Car] {engine};

    \draw[->,thick] (World) to (Car);
    \draw[->,thick] (World) to (Driver);
    \draw[->,thick] (Car) to (Engine);

    \draw[->,thick,dashed,cgreen] (Car) edge[bend left] node[right] {owner} (Engine);
    \draw[->,thick,dashed,cgreen] (Engine) edge[bend left] node[left] {inside} (Car);
    \draw[->,thick,dashed,cgreen] (Car) edge node[above] {sibling} (Driver);
    \draw[->,thick,dashed,red] (World) edge[bend right=90] node[left] {invalid} (Engine);
    \draw[->,thick,dashed,red] (Driver) edge[bend left] node[right] {invalid} (Engine);
  \end{tikzpicture}

  \caption{solid lines indicate "owns" and dotted lines indicate "references"}
\end{figure}
\end{frame}

\begin{frame}{Owners-as-dominators: implications}
\begin{itemize}
  \item any external reference to an object must go through its owner
  \item objects are only protected from external access, not internal access
  \item owners-as-dominators: owner is dominator for all of its owned objects
  \item doesn't allow common idioms that involve aliasing
  \item model restricts \emph{where} references can point, not \emph{how} references are used
\end{itemize}

\begin{figure}[fragile]
  \begin{tikzpicture}[object/.style={rectangle, draw, very thick, minimum size=20}]
    \node[object] (World) {world};
    \node[object] (Car) [below=of World] {car};
    \node[object] (Driver) [right=of Car,xshift=1.5cm] {driver};
    \node[object] (Engine) [below=of Car] {engine};

    \draw[->,thick] (World) to (Car);
    \draw[->,thick] (World) to (Driver);
    \draw[->,thick] (Car) to (Engine);

    \draw[->,thick,dashed,cgreen] (Car) edge[bend left] node[right] {owner} (Engine);
    \draw[->,thick,dashed,cgreen] (Engine) edge[bend left] node[left] {inside} (Car);
    \draw[->,thick,dashed,cgreen] (Car) edge node[above] {sibling} (Driver);
    \draw[->,thick,dashed,red] (World) edge[bend right=90] node[left] {invalid} (Engine);
    \draw[->,thick,dashed,red] (Driver) edge[bend left] node[right] {invalid} (Engine);
  \end{tikzpicture}
\end{figure}
\end{frame}

\begin{frame}{Owners-as-modifiers}
Same rules as owners-as-dominators, but \emph{a} can also reference \emph{b} through \emph{r}, if:
\begin{enumerate}
  \item \emph{r} is a read-only reference and only pure methods (that don't modify existing objects) can be called on it.
\end{enumerate}

\begin{figure}
  \begin{tikzpicture}[object/.style={rectangle, draw, very thick, minimum size=20}]
    \node[object] (World) {world};
    \node[object] (Car) [below=of World] {car};
    \node[object] (Driver) [right=of Car,xshift=1.5cm] {driver};
    \node[object] (Engine) [below=of Car] {engine};

    \draw[->,thick] (World) to (Car);
    \draw[->,thick] (World) to (Driver);
    \draw[->,thick] (Car) to (Engine);

    \draw[->,thick,dashed,cgreen] (Car) edge[bend left] node[right] {owner} (Engine);
    \draw[->,thick,dashed,cgreen] (Engine) edge[bend left] node[left] {inside} (Car);
    \draw[->,thick,dashed,cgreen] (Car) edge node[above] {sibling} (Driver);
    \draw[->,thick,dashed,orange] (World) edge[bend right=90] node[left] {read only} (Engine);
    \draw[->,thick,dashed,orange] (Driver) edge[bend left] node[right] {read only} (Engine);
  \end{tikzpicture}
\end{figure}
\end{frame}

\begin{frame}{Owners-as-modifiers: implications}
\begin{itemize}
  % \item any external reference to an object must go through its owner
  % \item objects are only protected from external access, not internal access
  \item owners-as-modifiers: only the owners can modify objects
  % \item doesn't allow common idioms that involve aliasing
  % \item model restricts \emph{where} references can point, not \emph{how} references are used
\end{itemize}

\begin{figure}
  \begin{tikzpicture}[object/.style={rectangle, draw, very thick, minimum size=20}]
    \node[object] (World) {world};
    \node[object] (Car) [below=of World] {car};
    \node[object] (Driver) [right=of Car,xshift=1.5cm] {driver};
    \node[object] (Engine) [below=of Car] {engine};

    \draw[->,thick] (World) to (Car);
    \draw[->,thick] (World) to (Driver);
    \draw[->,thick] (Car) to (Engine);

    \draw[->,thick,dashed,cgreen] (Car) edge[bend left] node[right] {owner} (Engine);
    \draw[->,thick,dashed,cgreen] (Engine) edge[bend left] node[left] {inside} (Car);
    \draw[->,thick,dashed,cgreen] (Car) edge node[above] {sibling} (Driver);
    \draw[->,thick,dashed,orange] (World) edge[bend right=90] node[left] {read only} (Engine);
    \draw[->,thick,dashed,orange] (Driver) edge[bend left] node[right] {read only} (Engine);
  \end{tikzpicture}
\end{figure}
\end{frame}

\begin{frame}{From theory to Rust}
\begin{itemize}
\end{itemize}
\end{frame}

\section{Ownership in Rust}

\begin{frame}{Memory allocation system}
\end{frame}

\begin{frame}{Transferring ownership}
\end{frame}

\begin{frame}{References and borrowing}
\end{frame}

\section{Evaluation: another implementation}

% \subsection{Another ownership implementation}

\section{Conclusions and outlook}

\begin{frame}
  Here are some conclusions.
\end{frame}

\end{document}